---
title: "Intro to basics"
description: In this chapter, you will take your first steps with R. You will learn
  how to use the console as a calculator and how to assign variables. You will also
  get to know the basic data types in R. Let's get started!
title_meta: Chapter 1
output: html_notebook
---

# Arithmetic with R

In its most basic form, R can be used as a simple calculator. Consider the following arithmetic operators:

- Addition: `+`
- Subtraction: `-`
- Multiplication: `*`
- Division: `/`
- Exponentiation: `^`
- Modulo: `%%`

The last two might need some explaining:

- The `^` operator raises the number to its left to the power of the number to its right: for example `3^2` is `9`.
- The modulo returns the remainder of the division of the number to the left by the number on its right, for example 5 modulo 3 or `5 %% 3` is 2.

With this knowledge, follow the instructions below to complete the exercise.

### Instructions
- Type `2^5` in the editor to calculate 2 to the power 5.
```{r}
2^3
```
- Type `28 %% 6` to calculate 28 modulo 6.
```{r}
28 %% 6
```

- Note how the `#` symbol is used to add comments on the R code.

### Hint
Another example of the modulo operator: `9 %% 2` equals `1`.
```{r, echo=TRUE}
# Another example of the modulo operator: `9 %% 2` equals `1`.
9 %% 2
```


# Variable assignment 

A basic concept in (statistical) programming is called a **variable**. 

A variable allows you to store a value (e.g. 4) or an object (e.g. a function description) in R. You can then later use this variable's name to easily access the value or the object that is stored within this variable. 

You can assign a value 4 to a variable `my_var` with the command

```{r}
my_var <- 4
```

## over to you: 
- Complete the code in the editor such that it assigns the value 42 to the variable `x` in the editor. Notice that when you ask R to print `x`, the value 42 appears.
```{r}

```

## hint
- Look at how the value 4 was assigned to `my_variable` in the exercise's assignment. Do the exact same thing in the editor, but now assign 42 to the variable `x`.

- Remember that if you want to assign a number or an object to a variable in R, you can make use of the assignment operator `<-`. Alternatively, you can use `=`, but `<-` is widely preferred in the R community.

```{r, echo=TRUE}
# Assign a value to the variables my_apples and my_oranges
my_apples  <- 5
my_oranges <- 6

# Add these two variables together
my_apples + my_oranges

# Create the variable my_fruit
my_fruit <- my_apples + my_oranges
```

## watch out!
Be careful with the operations between different types/classes of objects
```{r, echo=TRUE}
# Assign a value to the variable my_apples
my_apples <- 5 

# Fix the assignment of my_oranges
my_oranges <- "six" 
```

```{r, echo=TRUE}
# Create the variable my_fruit and print it out
my_fruit <- my_apples + my_oranges
```

```{r, echo=TRUE}
class(my_oranges)
```
```{r, echo=TRUE}
class(my_oranges)
```
So, in general, it's a good idea to check that the objetcs that are opperating between each other, are of the same class/type or we have to be conscients that sometimes, if the types are not equals but they are "almost operables", *R* will change at least one of them to a type that make both be "totaly operables". 

There could be some warnings about this... it could be a good idea to knoe a little more about the **data types** that will be jumping in at our work.



